---
title: "Week5CH12"
author: "Nitin"
date: "June 8, 2019"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}
#options(tinytex.verbose = TRUE)
```
#Tidy data
Required package
```{r}
library(tidyverse)
library(tinytex)
```

See table
```{r}
table1
```

See table2
```{r}
table2
```

See table3
```{r}
table3
```

See table4a
```{r}
table4a
```

See table4b
```{r}
table4b
```

Diffrent ways to work with table 1.
Compute rate per 10,000
```{r}
table1 %>% 
  mutate(rate = cases / population * 10000)
```

Compute cases per year
```{r}
table1 %>% 
  count(year, wt = cases)
```

Visualise changes over time
```{r}
library(ggplot2)
ggplot(table1, aes(year, cases)) + 
  geom_line(aes(group = country), colour = "grey50") + 
  geom_point(aes(colour = country))
```

## Spreading and gathering
### Gathering
some times in data the column names are not names of variables, but values of a variable. 
See table4
```{r}
table4a
```


Generate the call to gather()
To tidy a dataset like this, we need to gather those columns into a new pair of variables.
```{r}
table4a %>% 
  gather(`1999`, `2000`, key = "year", value = "cases")
```

Generate the call to gather()
```{r}
table4b %>% 
  gather(`1999`, `2000`, key = "year", value = "population")
```

Combine the tidied versions of table4a and table4b into a single tibble,
```{r}
tidy4a <- table4a %>% 
  gather(`1999`, `2000`, key = "year", value = "cases")
tidy4b <- table4b %>% 
  gather(`1999`, `2000`, key = "year", value = "population")
left_join(tidy4a, tidy4b)

```

### Separate
In spreading opposite to what is done gathering.
```{r}
table2
```

```{r}
table2 %>%
    spread(key = type, value = count)
```



## Separating and uniting
Separate

```{r}
table3
```

Function separate() pulls apart one column into multiple columns, by splitting wherever a separator character appears
```{r}
table3 %>% 
  separate(rate, into = c("cases", "population"))
```

rate column contains both cases and population variables which need to be split it into two variables. 
```{r}
table3 %>% 
  separate(rate, into = c("cases", "population"), sep = "/")
```

separate() split the values of rate at the forward slash characters
```{r}
table3 %>% 
  separate(rate, into = c("cases", "population"), convert = TRUE)
```

```{r}
table3 %>% 
  separate(year, into = c("century", "year"), sep = 2)
```

# Unite
```{r}
table5 %>% 
  unite(new, century, year)
```

```{r}
table5 %>% 
  unite(new, century, year, sep = "")
```

## Missing values
A value can be missing in one of two possible ways.
1. Explicitly
2. Implicitly
```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2015, 2015, 2016, 2016, 2016),
  qtr    = c(   1,    2,    3,    4,    2,    3,    4),
  return = c(1.88, 0.59, 0.35,   NA, 0.92, 0.17, 2.66)
)
```

Make the implicit missing value explicit by putting years in the columns
```{r}
stocks %>% 
  spread(year, return)
```

Turn explicit missing values implicit by seting na.rm = TRUE
```{r}
stocks %>% 
  spread(year, return) %>% 
  gather(year, return, `2015`:`2016`, na.rm = TRUE)
```

Make missing values explicit in tidy data by using function complete
```{r}
stocks %>% 
  complete(year, qtr)
```

```{r}
treatment <- tribble(
  ~ person,           ~ treatment, ~response,
  "Derrick Whitmore", 1,           7,
  NA,                 2,           10,
  NA,                 3,           9,
  "Katherine Burke",  1,           4
)
```

```{r}
treatment %>% 
  fill(person)
```

## Case Study
```{r}
who
```

Focus on the values that are present.
```{r}
who1 <- who %>% 
  gather(new_sp_m014:newrel_f65, key = "key", value = "cases", na.rm = TRUE)
who1
```

Count using key.
```{r}
who1 %>% 
  count(key)
```

Makes all variable names consistent.
```{r}
who2 <- who1 %>% 
  mutate(key = stringr::str_replace(key, "newrel", "new_rel"))
who2
```

Separate the values in each code with two passes usinf function separate()
```{r}
who3 <- who2 %>% 
  separate(key, c("new", "type", "sexage"), sep = "_")
who3
```

```{r}
who3 %>% 
  count(new)
```

```{r}
who4 <- who3 %>% 
  select(-new, -iso2, -iso3)
```

Separate sexage into sex and age by splitting after the first character.
```{r}
who5 <- who4 %>% 
  separate(sexage, c("sex", "age"), sep = 1)
who5
```

who dataset is now tidy!
```{r}
who %>%
  gather(key, value, new_sp_m014:newrel_f65, na.rm = TRUE) %>% 
  mutate(key = stringr::str_replace(key, "newrel", "new_rel")) %>%
  separate(key, c("new", "var", "sexage")) %>% 
  select(-new, -iso2, -iso3) %>% 
  separate(sexage, c("sex", "age"), sep = 1)
```



